---
title: "Tiny Project for BST 625"
author: "First Name Last name (XXXX@umiami.edu)"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## MTcars data
This dataset contains a subset of the fuel economy data that the EPA makes available on http://fueleconomy.gov. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car.

The variables are

Variable | Content
----------|---------
displ| engine displacement, in litres
cyl (grid)| number of cylinders
trans| type of transmission
drv (grid) | the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd
cty | city miles per gallon
hwy | highway miles per gallon
class | "type" of car

```{r}
#mfrow(c(1,2))
library(ggplot2) 
ggplot(data = mpg) +
  aes(x = displ, y = cty, colour = class) + 
  geom_point() +
  facet_grid(drv ~ cyl) +
  ggtitle("Plot of mpg data \n using ggplot2") +
  xlab("Engine displacement, in litres") + ylab("City miles per gallon") +
  theme(axis.text = element_text(size = 9),
               axis.title = element_text(size = 14, face = "bold"),
               plot.title = element_text(size = rel(2)),
               legend.title = element_text(face = "bold"),
               panel.background = element_rect(fill = "white", colour = "grey50"))  

```

    
```{r cars,fig.width = 6,fig.height = 4}
ggplot(data = mpg) +
  aes(x = displ, y = hwy) + 
  geom_point(aes(colour = cyl)) + 
  # Default smoother is LOESS
  geom_smooth() +
  geom_smooth(method = "lm", colour = "red") +
  xlab("Engine displacement, in litres") + ylab("highway miles per gallon") +
  theme(axis.text = element_text(size = 9),
               axis.title = element_text(size = 14, face = "bold"),
               plot.title = element_text(size = rel(2)),
               legend.title = element_text(face = "bold"))  
```



