---
title: 'BST 625: Assignment 1 (R portion, 5 points in total)'
author: "Your Name"
date: "Due date 10/21/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Question 1 (1 point)

First I wrote down 

```{r}
dat <- read.csv("http://courses.washington.edu/b517/Datasets/shoulder.csv")
```

then there is an object created in my environment. Can you guess what I am doing? (1 point) If your answer is no, you will still get your point since we have not covered this yet.

```{r}
## write down your answer after "## ---|" in the next line. Provide code as you will
## ----| You are creating a data.frame by reading in a CSV file. (Now we have learnt that if read_csv() is used from {tidyverse}, it will be a tibble)
head(dat)
```
# Question 2 (2 points)

Then I wrote

```{r}
Pain_long <- dat$pain
```

What is *Pain_long*? (2 points)

```{r echo = TRUE}
## write down your answer after "## ---| in the next line". Provide code as you will
## ----| Pain_long is a vector extracted from the column labeled "pain" in the data.frame dat
head(Pain_long)
```
# Question 3 (2 points)

Conduct some operation on *Pain_long*. In other words, provide any function using *Pain_long* as an input (1 point). Hint: a function is in the format of *functionname()*. Describe your operation in English  (1 point) 

```{r}
## Tell R what you want to do in the next line
head(Pain_long, 10)
## write down your answer after "## ---|" in the next line. Provide code as you will
## ----| This function just prints out the first 10 rows of data
```
# Bonus 

Share something with me about what you have learned by yourself **beyond** what we have covered so far (1 point)

```{r}
## Tell R your command in the next line
mean(Pain_long[which(dat$sex==1)])
mean(Pain_long[which(dat$sex==2)])
## write down the purpose of the above command
## ----| to calculate conditional mean of Pain_long according to different gender. The which function will return the position of the elements in a logical vector where it only selects what is TRUE. Therefore the first statement will only select males because it is true when they are equal to 1.
```
