---
title: 'BST 625: Assignment 2 (R portion, 10 points in total)'
author: "Your Name"
date: "Due date 11/06/2020"
output:
  word_document: default
  pdf_document: default
---
Use this .Rmd R markdown file to generate a pdf or word document to submit through blackboard

(1)	Open the .Rmd file in your R studio. When you opening it at the first time, R studio will remind you to install packages that you need on the top of your file and you should click yes and then your “knit” button will turn blue. If you missed the chance of clicking “yes”, you can close R studio and reopen it again

(2)	Instead of clicking “run”, you will click the “Knit” button to generate a pdf or word document in your working directory (how to set up working directory is in our slides “R1_Intro.ppt” page 4 where source file location means the folder where you keep your .Rmd file). There is a pointing down arrow near the “Knit” button where you can choose pdf or word (see the attachment in our blackboard under HW4 assignment item). If you click “knit” the first time and have an error, you may have to click “Knit” again
 
(3)	Type in R command as instructed, “Knit”, and submit the generated document through blackboard. If you are approaching the deadline and you are unable to generate your pdf or word document, you can type in your answers in the attached word document and submit it

More content of R Markdown can be found in our class slides “R1_RMarkdown.ppt” and "Min_R_Manual.pdf Chapter 7"


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

# Question 1 (1 point)
## Import data
**Import data "flights.csv" into your environment**. My command is written as follows. You have to paste your own command. 
```{r}
flights <- read_csv("C:/Users/mlu6/Dropbox/R book/R3/flights.csv")
```
Our example data set has measurements on all flights departing the Greater New York City area in 2013. It has 1,000 unique flights and 19 measurements on those flights. For more information, you can install the package *nycflights13*, use command *library(nycflights13)* and *?flights* to get more information from the help file. The orginal dataset has 336,776 rows but I use command *write.csv(flights[sort(sample(1:336776, 1000)), ], file = "flights.csv", row.names = F)* to sample only 1000 rows for our practice. We will use this data set to explore a few of the more commonly-used dplyr functions. 

# Question 2 (2 points)
## Subset Rows with filter()
**Inspect only flights that happened on December 24th**. The target columns are *month* and *day*. (Hint: the filter function takes in the data frame and a logical argument. Rows that return TRUE for that logical argument are returned as a subset of the original data frame. For example, when I wrote *filter(flights, year == 2013, month == 1)* or *flights %>% filter(year == 2013, month == 1)*, I inspected flights that happened in January, 2013).
```{r}
## Tell R what you want to do in the next line

```
# Question 3 (2 points)
## Sort Rows with arrange()
**Sort the flights by departure delay time in decreasing order**. The target column is *dep_delay* and the target function for decreasing order is *desc()*. (Hint: The *arrange()* function takes in a data frame and a set of columns to sort on. We can sort the flights by in increasing order by the day of the year using command *arrange(flights, year, month, day)*: we sort by month, then by day within month. If we want to sort month in a decrease order, we can use *arrange(flights, desc(month))*).
```{r echo = TRUE}
## Tell R what you want to do in the next line

```

# Question 4 (2 point)
## Pipes command with %>%
I wrote the following command. Note that all of these compose one command.
```{r}
flights %>%
  filter(carrier == "AA") %>% 
  filter(origin == "JFK") %>% 
  filter(month == 11) %>% 
  arrange(desc(air_time)) 
```
**What I am inspecting?** (1 point)
```{r}
## write down the purpose of the above command
## ----|
```
**Can you modify the above command to inspect something else?** (1 point)
```{r echo = TRUE}
## Tell R what you want to do in the next line

```

# Question 5 (3 points)
## Select and Move Columns with select()
**Find all the columns with names starting with “dep_”**.(2 points) The target function for starting with specific character is *starts_with()* (Hint: the command for finding all the columns with names ending with “ay” is *select(flights, ends_with("ay"))*).
```{r}
## Tell R what you want to do in the next line

```
I want to add an indicator for “long flights” (American Airlines considers a long flight any flight over four-and-a-half hours) and then order these flights by flight time, excluding flights to Honolulu. I also want to rearrange the columns *dest*, *long_flight*, *air_time* up to front. I wrote
```{r}
flights %>%
  mutate(long_flight = air_time >= 270) %>% 
  filter(dest != "HNL") %>% 
  select(dest, long_flight, air_time, everything()) %>% 
  arrange(desc(air_time)) 
```
**Can you modify the above command to add an indicator for "short flights" and then order these flights by departure delay, excluding flights that depart from Newark Liberty?** Target columns are *dep_delay* and *origin*. (Hint: you need argument *origin != "EWR"*)
```{r}
## Tell R what you want to do in the next line

```

# Bonus (10 points)

Share something with me about what you have learned by yourself **beyond** what we have covered so far (1 point)
```{r}
## Tell R your command in the next line

## write down the purpose of the above command
## ----|
```
**Read "3-data-manipulat_ProjectExample1.R" and use the data in your Assignment 1 and create a table like this (9 points: 3 points for each step as follows or you can have your own steps)**

Age group | Control |Treatment 
----------|---------|----
24- |        1 |        1
25-34|       3  |       2
35-44|       2  |       3
45-54|       1  |       4
55-64|       4  |       6
65-74|       7  |       2
75+ |        4  |       1

Hint: **First**, I reshaped the data from long format to wide format:
```{r}
#dat <- read_csv("http://courses.washington.edu/b517/Datasets/shoulder.csv")
#dat <- spread(dat, key = time, value = pain)
```
For {base} users, the command is 
```{r}
#dat <- read.csv("http://courses.washington.edu/b517/Datasets/shoulder.csv")
#dat <- reshape(dat, v.names = "pain", idvar = "id",
#               timevar = "time", direction = "wide")
```
**Second**, I used *trt* and *age* to create two new columns *age_grp* and *ct_grp*. The data looked like this

id | trt | sex | age | pain.1 | pain.2 | pain.3 | pain.4 | pain.5 | pain.6 | age_grp | ct_grp
---|-----|-----|-----|--------|--------|--------|--------|--------|--------|--------|------
101 |  1 |  2 | 64   |   0    |  0     | 0      |0      |0      |0|   55-64 |  control
102 |  1 |  1 | 41   |   2    |  1     | 0      |0      |0      |0|   35-44 |  control
103 |  1 |  2 | 77   |   2    |  1     | 1      |1      |0      |0|     75+ |  control
104 |  1 |  2 | 54   |   0    |  0     | 0      |0      |0      |0|   45-54 |  control


For {tidyverse} users,  you have to use command like
```{r}
#df <- df %>% 
#  mutate(
#    age_grp = 
#      case_when(
#        age == 1 ~ "25-34", 
#        age == 2 ~ "35-44", 
#        age == 3 ~ "45-54", 
#        age == 4 ~ "55-64", 
#        age == 5 ~ "65-74", 
#        TRUE ~ "75+", 
#      )
#  )
```
But instead of using "age == 1", you have to use something like *age < 25* and *(25 <= age)&(age < 35)*.

For {base} users, you can use the cut() function to create a factor. For adding a new column, you can directly use the assignment sign "dat$newcolumnName <-"
```{r}
#dat$age_grp <- cut(dat$age, breaks = c(0, 24, 34, 44, ....
#                   labels = c("24-", "25-34", "35-44", "45-54"....
```
**Third**, I create a summary table.

For {tidyverse} users,  you have to use command like
```{r}
#dat %>% 
#  group_by(age_grp, ct_grp) %>% 
#  summarise(n = n()) %>% 
#  spread(ct_grp, n) 
```

For {base} users,  you can use command like
```{r}
#table(dat$age_grp,dat$ct_grp) 
```
**Paste all your commands here**
```{r}
## Tell R your commands in this block

```