---
title: 'BST 625: Assignment 3 (R portion, 15 points in total)'
author: "Your Name"
date: "Due date 11/23/2020"
output:
  word_document: default
  pdf_document: default
---
Use this .Rmd R markdown file to generate a pdf or word document to submit through blackboard

# Our goal is to conduct a case study and you do not have to do the {base} user part at all. If you do the {base} user part, there won't be any extra points earned but I can provide feedback to you if any. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```


# Import and reshape data (long to wide)
```{r}
dat <- read_csv("http://courses.washington.edu/b517/Datasets/shoulder.csv")
dat
```
In laparoscopic gallbladder surgery, carbon dioxide gas is forced into the 
abdomen to separate the organs and improve visibility and access to the 
gallbladder.  One side-effect of this is to separate the liver from the 
diaphragm, where it is usually supported in part by surface tension.  The 
resulting strain on the tissues attaching the two causes pain that is felt 
in the shoulder (`referred pain').  This can be more severe than the 
pain from the (relatively minor) surgical incisions.

A randomised trial explored using suction to drain out the carbon dioxide 
gas.  Pain in the shoulder, at the incision, and over the abdomen was 
assessed twice daily for three days by a questionnaire.  There was no sign 
of any difference in wound or abdominal pain.  You have data on shoulder 
pain, which has been recoded to a 0-4 scale, where 0 is no pain, and 4 is 
the most severe category reported.  The goal is to assess the 
effectiveness of this intervention.

The variables are

Variable | Content
----------|---------
id|  a patient identifier
trt|  1 is control, 2 is post-operative suction (treatment)
sex| 1 is male, 2 is female
age| in years
pain| 0-4 scale
time| 1 is the first, 6 is the last measurement on each person


Now let me reshape the data. Note that both *pivot_wider()* and *spread()* are from *{tidyverse}*: 
*pivot_wider()* is a new version of *spread()*.
```{r}
dat <- dat %>% 
  pivot_wider(names_from = time, 
              values_from = pain, 
              names_prefix = "pain")
dat
```
# Question 1: create a new variable *Avg_Pain* (1 point)
**Create a new variable Avg_pain in dat as the mean of pain1 to pain6**. The target columns are *pain1* to *pain6*. (Hint: find the functions that you like in "R5_SAS_project 2_3_in_R.ppt". For {tidyverse} user, you can use *mutate* and *rowMeans(select(.,pain1:pain6)*, instead of inputing *pain1, pain2, pain3, pain4, pain5 and pain6* one by one. For {base} user, *rowMeans(dat[,XXXX])* is also used: you can use *paste("pain", 1:6, sep = "")* to replace *c("pain1", "pain2", "pain3", "pain4", "pain5", "pain6")*). DON'T FORGET THE ASSIGNMENT SIGN *<-*.
```{r}
## Tell R your command in the next line




```

```{r}
## If you want to try {base}, tell R your command in the next line
dat$Avg_Pain <- rowMeans(dat[,paste("pain", 1:6, sep = "")], na.rm = TRUE)
```

# Question 2: subset your dataset and write out a .csv file (4 points)
**Create a file, Q1.pain_m.csv (1 point but you don't have to upload this .csv file), where there are 4 variables only --- {id, trt, age, Avg_Pain} (1 point). This file contians male patients only (1 point) and Avg_Pain is sorted from high to low (1 point).** The target columns for column selection are *id, trt, age* and *Avg_pain*  and the target columns for row manipulation are *sex* and *Avg_Pain*. (Hint: find the functions in "R5_SAS_project 2_3_in_R.ppt". For {base} user, you can use *x <- dat[which(dat$sex == 1), c("id", "trt", "age", "Avg_Pain")]* and *write.csv(x[order(x$Avg_Pain, decreasing = TRUE),], file = "Q1.pain_m.csv", row.names = FALSE)*).
```{r}
## Tell R your {tidyverse} command in the next line






```

```{r}
## If you want to try {base}, tell R your command in the next line

```
# Question 3: summarize your dataset by group (1 point)
I wrote the following command in both {tidyverse} and {base} format.
```{r}
## Tell R your {tidyverse} command in the next line
  dat %>% 
  group_by(sex,trt) %>% 
  count(pain6) %>%
  spread(pain6,n) 
```

```{r}
## Tell R your {base} command in the next line
ftable(dat[, c("sex", "trt", "pain6")])
```
**What I am inspecting?(0.5 point)** 
```{r}
## write down the purpose of the above command
## ----|
```
**Can you modify the above command to inspect treatment and Avg_pain?(0.5 points)** The target columns are *trt* and *Avg_pain* and you have to round *Avg_pain* to integers to create tables. (Hint: for {tidyverse} user, instead of using *group_by(sex,trt)*, you should use *group_by(trt)*; instead of using *count(XXXX)*, you should use *count(round(XXXX))*. For {base} user, I suggest you use *table()* instead of *ftable()* since *ftable()* is more useful for more than 2 columns: here we only summarize two columns so you can use *table(dat$XXXX, round(dat$XXXX))*).
```{r}
## Tell R what you want to do in the next line





```



```{r}
## If you want to try {base}, tell R your command in the next line


```
# Question 4: function usage (1 point)
Let me fake a categorical variable of bladder type and fake a datset with missing values using {base} and rearange my columns of this new dataset using {tidyverse} for our task later.
```{r}
## Min's job
dat$bladder_type <- sample(c("A", "B", "C"), nrow(dat), replace = TRUE)
dat_miss <- dat
dat_miss[c(1,3,5), c("Avg_Pain", "bladder_type")] <- NA
dat_miss <- dat_miss %>% 
                  select(sex, age, Avg_Pain, bladder_type, pain1:pain3)
dat_miss
```
Now I create a function that when the missingness occurs in numeric columns, I will impute the mean of corresponding column and when the missingness occurs in categorical columns, I will impute the category that has the largest frequency. Again, learning {base} is very useful for creating your own function.
```{r}
## Min's job
na.imp <- function(x){
  if (is.numeric(x)){
  x[is.na(x)] <- mean(x, na.rm = TRUE)
  } else {
  tb <- table(x)
  x[is.na(x)] <-  names(tb)[which.max(tb)[1]] ##  [1] is to choose the first one for ties
  }
  x
}
```

**Can you use map() and the function above function na.imp() to impute the missingness in all the columns in dat_miss?(1 point)**  Target slide is *R5_Explor_data_analy_Own_func.ppt # 18*.  (Hint: you have use *as_tibble* at the end of your pipeline to convert the returned list from *map()* to tibble. For {base} user, you can use *as.data.frame(lapply(dat_miss, na.imp))*.)
```{r}
## Tell R what you want to do in the next line





```


```{r}
## If you want to try {base}, tell R your command in the next line


```


# Question 5: data visulization (7 points)
## 5.0 data preparation (1 point)
**Use the solution of HW5 bonus part and create two new columns in dat with grouped age and recoded treatment group** Hint: the target slide is *R5_SAS_project 2_3_in_R.ppt # 6*, where you do not need to put rowwise() since rowwise() for calculating sum in that slide.Your updated *dat* should look like this


id | trt | sex | age | pain.1 | pain.2 | pain.3 | pain.4 | pain.5 | pain.6 | age_grp | ct_grp
---|-----|-----|-----|--------|--------|--------|--------|--------|--------|--------|------
101 |  1 |  2 | 64   |   0    |  0     | 0      |0      |0      |0|   55-64 |  control
102 |  1 |  1 | 41   |   2    |  1     | 0      |0      |0      |0|   35-44 |  control
103 |  1 |  2 | 77   |   2    |  1     | 1      |1      |0      |0|     75+ |  control
104 |  1 |  2 | 54   |   0    |  0     | 0      |0      |0      |0|   45-54 |  control



```{r}
## Tell R your command in the next line











```

## 5.1 barplot (1.5 points)
**Can you generate a barplot that looks like this (check out the word document)?** Hint: the target function is *ggplot(dat) + geom_bar(mapping = aes(x = XXXX)) + labs(x = "XXXX", y="XXXX")* and ususally we put whatever after *+* in the next line to make our command look neat. 

```{r, fig.width = 5,fig.height = 3}



```

## 5.2 dotplot (1.5 points)
**Can you generate a dot plot that looks like this (check out the word document)?** Hint: the target function is *ggplot(dat) + geom_count(mapping = aes(x = XXXX, y = XXXX)) + labs(x = "XXXX", y="XXXX")* and ususally we put whatever after *+* in the next line to make our command look neat.  
```{r, fig.width = 5,fig.height = 3}
## Tell R your {base} command in the next line




```

## 5.3 mapplot (1.5 points)
**Can you generate a dot plot that looks like this (check out the word document)?**  
Hint: the target function is *dat %>% count(XXXX, XXXX) %>% ggplot(mapping = aes(x = XXXX, y = XXXX)) + geom_tile(mapping = aes(fill = n))+ labs(x = "XXXX", y="XXXX")* and ususally we put whatever after *%>%* and *+* in the next line to make our command look neat.
```{r, fig.width = 5,fig.height = 3}
## Tell R your {base} command in the next line




```


## 5.4 scatterplot (1.5 points)
I made the following two plots.
```{r, fig.width=5,fig.height=3}
## Tell R your {base} command in the next line
ggplot(dat, aes(x=age, y=Avg_Pain)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm)   # Add linear regression line 
                             #  (by default includes 95% confidence region)
```

```{r, fig.width=5,fig.height=3}
## Tell R your {base} command in the next line
ggplot(dat, aes(x=age, y=Avg_Pain)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth()            # Add a loess smoothed fit curve with confidence region
                             #> `geom_smooth()` using method = 'loess'
```

**What is the difference between the two plots (0.5 point)** 

```{r}
## write down your answer in the next line after ## ----|
## ----|
```

**Can you make some modifications on the above two plots?(1 point)** (Hint: my purpose for this question is only for demonstration since I never explain confidence region at all. Teeny-tiny modification will be enough for this question.)
```{r}
## Tell R your commands in the next line


```
# Question 5: pipeline connection (1 point)
**Can you connect your commands in Question 1, Question 5.0 and Question 5.1 using pipe operator %>%?** 
```{r}
## Tell R your commands in the next line











```
# Bonus (2.5 points)
## Self-learning ability (1 point)
Share something with me about what you have learned by yourself **beyond** what we have covered so far (1 point)
```{r}
## Tell R your command in the next line

## write down the purpose of the above command
## ----|
```

## {base} plot (1.5 point)
I wrote the following command to explore the relationship among gender, treatment and my outcome.
```{r}
cols <- rainbow(2, s = 0.5)
boxplot(Avg_Pain ~ trt + sex, data = dat,
          at = c(1:2 , 4:5) , col = cols ,
          xaxs = FALSE)
```

**Can you modify the above command to make a better plot?(1 point)** 

Hint: you can change the *XXXX* part in the next block and you can go to my *Min_R_Manual.pdf #18-19* to use colors directly such as *cols <- c("chocolate1", "chartreuse4")* --- all these colors are from {base} R and you can also use numbers to make assignment for colors such as  *cols <- 1:2* for a quick plot.
```{r}
## Tell R your command in the next line
cols <- rainbow(2, s = 0.5)
boxplot(Avg_Pain ~ trt + sex, data = dat,
          at = c(1:2 , 4:5) , col = cols ,
          names = c("XXXXA", "XXXXB", "XXXXA", "XXXXB"),
          xaxs = FALSE,
        xlab = "XXXXC",
        ylab = "XXXXD")
legend("topleft", fill = cols ,
          legend = c("XXXX1", "XXXX2") , horiz = T, bty = "n")
```

**What can the argument -- bty = "n" -- above do?(0.5 point)** 

```{r}
## write down the purpose of the above command
## ----|
```